Critical assessment of Linear-Response Excitations in Molecular system\\ or \\ On the Localization of the Linear-Response excitations in Open Systems

### 2018-01-25 LRDraft.tex
update

This formulation of the Liouville equation is the starting point for the derivation of the spectral decomposition of the response density operator
in terms of the eigenvectors if the Liouville operator. 
The first step in the achievement of this result is provided by the construction of the eigenvectors of the matrix \eqref{LiouvilleMatrixTransBasisDef1}.
Following the notation previously defined we introduce the rank-2 vectors $\sket{p\alpha}{E}$ and $\sbra{p\alpha}{\tilde E}$ that represent, by definition,
the right and left eigenvectors $\optr{L}$, respectively, so that:
\be\lb{LiouvMatrixEigenvectorDef1}
\sum_q \sint \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q \beta}{E} = E \sket{p\alpha}{E} \;, \quad
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \optr{L}^{p \alpha}_{q\beta} = E \sbra{q\beta}{\tilde E}\;.
\ee
In general, left and right eigenvectors do not coincide, but it is always possible to choose they normalization so that a orthonormality conditions among the
elements of left and right spaces holds, namely:
\be\lb{ExcitationOrthonormalityDef1}
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E}\sket{p\alpha}{E'} \equiv \dbraket{\tilde E}{E'} = \delta(E-E') \;.
\ee
Once that the eigenvectors are known, the spectral decomposition of the resolvent of $\optr{L}$ can be provided:  
\be
\left[\left(\bbomega-\optr{L}\right)^{-1}\right]^{p\alpha}_{q\beta} =\sint\dd E
\frac{\sket{p\alpha}{E}\sbra{q\beta}{\tilde E}}{\bbomega-E} \;,
\ee
so equation \eqref{LiouvilleTransBasisDef1} can be inverted:
\be\lb{RhopExcitationTransBasisDef1}
\sket{p\alpha}{\dm'(\omega)} = \sint\dd E \sket{p\alpha}{E}
\frac{\sum_q \sint \dd \beta\sbra{q\beta}{\tilde E}
\sket{q\beta}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;,
\ee
that can be compactly rewritten by using the index-less notation:
% \be\lb{RhopExcitationTransBasisDef2}
% \sket{p\alpha}{\dm'(\omega)} = 
% \sint\dd E \sket{p\alpha}{E}
% \frac{\dbraket{\tilde E}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;.
% \ee
\be\lb{RhopExcitationTransBasisDef2}
\dket{\dm'(\omega)} = 
\sint\dd E \dket{E}
\frac{\dbraket{\tilde E}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;.
\ee

@@

We introduce the explicit representation of the eigenvectors op $\optr{L}$:
\be\lb{XYcoeffDef1}
\sket{p\alpha}{E} \equiv \mat{X_E^{p\alpha} \\ Y_E^{p\alpha}} \;, \qq
\sbra{p\alpha}{\tilde E} \equiv \mat{\tilde X_{E\,p\alpha} & \tilde Y_{E\,p\alpha}} \;,
\ee
and on the basis of these elements we define the \emph{excitation operators}:
\begin{align}\lb{ExctitationOperatorsDef1}
\op E &\equiv \dbraket{\cdot}{E} = 
\sum_p \sint \dd \alpha \left(\excite{p}{\alpha} X_E^{p\alpha} + \decay{\alpha}{p} Y_E^{p\alpha}\right) 
\;,\nn \\
\op{\tilde E} &\equiv \dbraket{\tilde E}{\cdot} = \sum_p \sint \dd \alpha \left(\decay{\alpha}{p} \tilde X_{E\,p\alpha} + 
\excite{p}{\alpha}\tilde Y_{E\,p\alpha}  \right)\;.
\end{align}
The orthonormality condition \eqref{ExcitationOrthonormalityDef1} implies that:
\be
\trace{\op{\tilde E} \op E'} = \sum_p\sint \dd\alpha \left(\tilde X_E^{p\alpha} X_{E'\,p\alpha} +
\tilde Y_E^{p\alpha} Y_{E'\,p\alpha}\right) = \delta(E-E') \;. 
\ee
Excitation operators are the right and left eigenvectors of the Liouville superoperator, as directly follows from their
definition. Indeed for $\op E$ we have:
\begin{align}\lb{opEeigenvalueDef1}
\Liouv \op E & = \sum_p \sint \dd \alpha \left(\trace{\decay{\alpha}{p} \Liouv \op E} \excite{p}{\alpha} + 
\decay{\alpha}{p} \trace{ \excite{p}{\alpha} \Liouv \op E}  \right) = \nn \\
&= \sum_p \sint  \dd \alpha \sbra{p\alpha}{\cdot} \sum_q \sint  \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q\beta}{E} 
= E \sum_p \sint \dd \alpha \sbra{p\alpha}{\cdot} \sket{p\alpha}{E} = E \, \op E \;,	
\end{align}
whereas the left action on $\op{\tilde E}$ provides:
\begin{align}\lb{optildeEeigenvalueDef1}
 \op{\tilde E} \Liouv &=  \sum_p \sint \dd \alpha \left(
\trace{\op{\tilde E}\Liouv\excite{p}{\alpha}}\decay{\alpha}{p} +  
\trace{\op{\tilde E}\sop{\mathcal L}\decay{\alpha}{p}} \excite{p}{\alpha} \right) = 
\nn \\ &= \sum_p \sint  \dd \alpha  \sum_q \sint  \dd \beta
\sbra{q\beta}{\tilde E}\optr{L}^{q\beta}_{p\alpha}\sket{p\alpha}{\cdot} = 
E\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \sket{p\alpha}{\cdot}  =
E \op{\tilde E} \;.
\end{align}
The introduction of the excitation operators allows to provide a complete operator version of equation \eqref{RhopExcitationTransBasisDef2}, that is:
\be\lb{RhopExcitationDef1}
\dm'(\omega) = 
\sint\dd E \op E
\frac{\trace{\op{\tilde E}\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;,
\ee
where the algebraic equivalence between the trace written in \eqref{RhopExcitationDef1} and the numerator of \eqref{RhopExcitationTransBasisDef2}
can be easily proved.

\subsubsection{Sternheimer equation for excitonic states. Localization argument revisited}

We introduce the \emph{excitonic states} defined via the linear combination of the basis set orthogonal to $\dmnot$ with the
coefficients that appears in \eqref{XYcoeffDef1}, that is:
\begin{align}\lb{ExcitonicStatesDef1}
\ket{\phi^E_p} &\equiv \sint \dd \alpha X_{E}^{p \alpha} \ket{\psi(\alpha)} &
\tbra{\phi^E_p} &\equiv \sint \dd \alpha {\tilde X_{E}^{p \alpha}} \bra{\psi(\alpha)} \;, \nn\\
\bra{\chi^E_p} &\equiv \sint \dd \alpha  Y_{E}^{p \alpha} \bra{\psi(\alpha)} &
\tket{\chi^E_p} &\equiv \sint \dd \alpha {\tilde Y}_{E}^{p\alpha} \ket{\psi(\alpha)} \;.
\end{align}
Thanks to these definitions we can express the excitation operators as follows:
\be\lb{ExctitationOperatorsDef2}
\op E = \sum_p \left( \ket{\phi^E_p}\bra{\psi_p} + \ket{\psi_p} \bra{\chi^E_p}\right)\;, \qq
\op {\tilde E} = \sum_p \left( \ket{\psi_p} \tbra{\phi^E_p} + \tket{\chi^E_p}\bra{\psi_p} \right)\;.
\ee
The operator eigenvalue equations \eqref{opEeigenvalueDef1} and \eqref{optildeEeigenvalueDef1} can be translated into a Schroedinger-like equations
for the excitonic states.

>>>>

This formulation of the Liouville equation is the starting point for the derivation of the spectral decomposition of the response density operator
in terms of the eigenvectors if the Liouville operator. 
The first step in the achievement of this result is provided by the construction of the eigenvectors of the matrix \eqref{LiouvilleMatrixTransBasisDef1}.
Following the notation previously defined we introduce the rank-2 vectors $\sket{p\alpha}{E}$ and $\sbra{p\alpha}{\tilde E}$ that represent, by definition,
the right and left eigenvectors $\optr{L}$, respectively, so that:
\be\lb{LiouvMatrixEigenvectorDef1}
\sum_q \sint \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q \beta}{E} = E \sket{p\alpha}{E} \;, \quad
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \optr{L}^{p \alpha}_{q\beta} = E \sbra{q\beta}{\tilde E}\;.
\ee
In general, left and right eigenvectors do not coincide, but it is always possible to choose they normalization so that a orthonormality conditions among the
elements of left and right spaces holds, namely:
\be\lb{ExcitationOrthonormalityDef1}
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E}\sket{p\alpha}{E'} \equiv \dbraket{\tilde E}{E'} = \delta(E-E') \;.
\ee
Once that the eigenvectors are known, the spectral decomposition of the resolvent of $\optr{L}$ can be provided:  
\be
\left[\left(\bbomega-\optr{L}\right)^{-1}\right]^{p\alpha}_{q\beta} =\sint\dd E
\frac{\sket{p\alpha}{E}\sbra{q\beta}{\tilde E}}{\bbomega-E} \;,
\ee
so equation \eqref{LiouvilleTransBasisDef1} can be inverted and we obtain the response density written in basis of transistion, in the form:
% \be\lb{RhopExcitationTransBasisDef1}
% \sket{p\alpha}{\dm'(\omega)} = \sint\dd E \sket{p\alpha}{E}
% \frac{\sum_q \sint \dd \beta\sbra{q\beta}{\tilde E}
% \sket{q\beta}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;,
% \ee
% that can be compactly rewritten by using the index-less notation:
\be\lb{RhopExcitationTransBasisDef1}
\sket{p\alpha}{\dm'(\omega)} = 
\sint\dd E \sket{p\alpha}{E}
\frac{\dbraket{\tilde E}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;.
\ee
% \be\lb{RhopExcitationTransBasisDef2}
% \dket{\dm'(\omega)} = 
% \sint\dd E \dket{E}
% \frac{\dbraket{\tilde E}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;.
% \ee

@@

We introduce the explicit representation of the eigenvectors op $\optr{L}$:
\be\lb{XYcoeffDef1}
\sket{p\alpha}{E} \equiv \mat{X_E^{p\alpha} \\ Y_E^{p\alpha}} \;, \qq
\sbra{p\alpha}{\tilde E} \equiv \mat{\tilde X_{E\,p\alpha} & \tilde Y_{E\,p\alpha}} \;,
\ee
and on the basis of these elements we define the \emph{excitation operators}:
\begin{align}\lb{ExctitationOperatorsDef1}
\op E &\equiv \dbraket{\cdot}{E} = 
\sum_p \sint \dd \alpha \left(\excite{p}{\alpha} X_E^{p\alpha} + \decay{\alpha}{p} Y_E^{p\alpha}\right) 
\;,\nn \\
\op{\tilde E} &\equiv \dbraket{\tilde E}{\cdot} = \sum_p \sint \dd \alpha \left(\decay{\alpha}{p} \tilde X_{E\,p\alpha} + 
\excite{p}{\alpha}\tilde Y_{E\,p\alpha}  \right)\;.
\end{align}
The orthonormality condition \eqref{ExcitationOrthonormalityDef1} implies that:
\be
\trace{\op{\tilde E} \op E'} = \sum_p\sint \dd\alpha \left(\tilde X_E^{p\alpha} X_{E'\,p\alpha} +
\tilde Y_E^{p\alpha} Y_{E'\,p\alpha}\right) = \delta(E-E') \;. 
\ee
Excitation operators are the right and left eigenvectors of the Liouville superoperator, as directly follows from their
definition. Indeed for $\op E$ we have:
\begin{align}\lb{opEeigenvalueDef1}
\Liouv \op E & = \sum_p \sint \dd \alpha \left(\trace{\decay{\alpha}{p} \Liouv \op E} \excite{p}{\alpha} + 
\decay{\alpha}{p} \trace{ \excite{p}{\alpha} \Liouv \op E}  \right) = \nn \\
&= \sum_p \sint  \dd \alpha \sbra{p\alpha}{\cdot} \sum_q \sint  \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q\beta}{E} 
= E \sum_p \sint \dd \alpha \sbra{p\alpha}{\cdot} \sket{p\alpha}{E} = E \, \op E \;,	
\end{align}
whereas the left action on $\op{\tilde E}$ provides:
\begin{align}\lb{optildeEeigenvalueDef1}
 \op{\tilde E} \Liouv &=  \sum_p \sint \dd \alpha \left(
\trace{\op{\tilde E}\Liouv\excite{p}{\alpha}}\decay{\alpha}{p} +  
\trace{\op{\tilde E}\sop{\mathcal L}\decay{\alpha}{p}} \excite{p}{\alpha} \right) = 
\nn \\ &= \sum_p \sint  \dd \alpha  \sum_q \sint  \dd \beta
\sbra{q\beta}{\tilde E}\optr{L}^{q\beta}_{p\alpha}\sket{p\alpha}{\cdot} = 
E\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \sket{p\alpha}{\cdot}  =
E \op{\tilde E} \;.
\end{align}
The introduction of the excitation operators allows to provide a complete operator version of equation \eqref{RhopExcitationTransBasisDef1}, that is:
\be\lb{RhopExcitationDef1}
\dm'(\omega) = 
\sint\dd E \op E
\frac{\trace{\op{\tilde E}\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;,
\ee
where the algebraic equivalence between the trace written in \eqref{RhopExcitationDef1} and the numerator of \eqref{RhopExcitationTransBasisDef1}
can be easily proved.

\subsubsection{Excitation expansion of the fluctuation states. A localization argument for the excited states}

We introduce the \emph{excited states} defined via the linear combination of the basis set orthogonal to $\dmnot$ with the
coefficients that appears in \eqref{XYcoeffDef1}, that is:
\begin{align}\lb{ExcitonicStatesDef1}
\ket{\phi^E_p} &\equiv \sint \dd \alpha X_{E}^{p \alpha} \ket{\psi(\alpha)} &
\tbra{\phi^E_p} &\equiv \sint \dd \alpha {\tilde X_{E}^{p \alpha}} \bra{\psi(\alpha)} \;, \nn\\
\bra{\chi^E_p} &\equiv \sint \dd \alpha  Y_{E}^{p \alpha} \bra{\psi(\alpha)} &
\tket{\chi^E_p} &\equiv \sint \dd \alpha {\tilde Y}_{E}^{p\alpha} \ket{\psi(\alpha)} \;.
\end{align}
Thanks to these definitions we can express the excitation operators as follows:
\be\lb{ExctitationOperatorsDef2}
\op E = \sum_p \left( \ket{\phi^E_p}\bra{\psi_p} + \ket{\psi_p} \bra{\chi^E_p}\right)\;, \qq
\op {\tilde E} = \sum_p \left( \ket{\psi_p} \tbra{\phi^E_p} + \tket{\chi^E_p}\bra{\psi_p} \right)\;.
\ee
Starting from these equation we can derive the excitation expansion of the fluctuation states. Indeed an inspection of formula \eqref{rhoPrimeFluctuationStateDef1}
show that fluctuation states can be extracted from the response density operator as follows:
\be
\ket{f_p(\omega)} = \dm'(\omega)\ket{\psi_p} \;, \qq
\bra{f_p(-\omega)} = \bra{\psi_p}\dm'(\omega) \;, \nn
\ee
so plugging equations \eqref{ExctitationOperatorsDef2} in the formula \eqref{RhopExcitationDef1} and performing these operations we obtain:
\be\lb{FluctStateExcitationDef1}
\ket{f_p(\omega)} = \sint\dd E \frac{\trace{\op{\tilde E}\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \ket{\phi^E_p} \;, \qq
\bra{f_p(-\omega)} = \sint\dd E \bra{\chi^E_p}\frac{\trace{\op{\tilde E}\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} 
\ee

\subsubsection{Excitation expansion of the linear response functional}

We have that:
\be
\sket{p\alpha}{\commutator{\op\Phi(\omega)}{\dmnot}} = \sum_q\sint\dd\beta\, \optr{J}^{p\alpha}_{q\beta}
\sket{q\beta}{\op\Phi(\omega)} \;\;, \;{\rm with} \;\;
\optr{J}^{p\alpha}_{q\beta} = \delta_{pq}\delta(\alpha-\beta)\mat{1 & 0 \\ 0 & -1}
\ee
and the linear response functional \eqref{LinearResponseFunctDef1} can be written as:
\be
{\cal R}_{\op O}(\omega) = 
\sint\dd E \dbraket{\op O}{E}
\frac{\dbra{\tilde E}\optr{J}\dket{\op\Phi(\omega)}}{\bbomega-E} \;.
\ee

%\subsubsection{Sternheimer equation for excitonic states. Localization argument revisited}
\subsubsection{Sternheimer equation for excited states. Localization argument revisited}

% We introduce the \emph{excitonic states} defined via the linear combination of the basis set orthogonal to $\dmnot$ with the
% coefficients that appears in \eqref{XYcoeffDef1}, that is:
% \begin{align}\lb{ExcitonicStatesDef1}
% \ket{\phi^E_p} &\equiv \sint \dd \alpha X_{E}^{p \alpha} \ket{\psi(\alpha)} &
% \tbra{\phi^E_p} &\equiv \sint \dd \alpha {\tilde X_{E}^{p \alpha}} \bra{\psi(\alpha)} \;, \nn\\
% \bra{\chi^E_p} &\equiv \sint \dd \alpha  Y_{E}^{p \alpha} \bra{\psi(\alpha)} &
% \tket{\chi^E_p} &\equiv \sint \dd \alpha {\tilde Y}_{E}^{p\alpha} \ket{\psi(\alpha)} \;.
% \end{align}
% Thanks to these definitions we can express the excitation operators as follows:
% \be\lb{ExctitationOperatorsDef2}
% \op E = \sum_p \left( \ket{\phi^E_p}\bra{\psi_p} + \ket{\psi_p} \bra{\chi^E_p}\right)\;, \qq
% \op {\tilde E} = \sum_p \left( \ket{\psi_p} \tbra{\phi^E_p} + \tket{\chi^E_p}\bra{\psi_p} \right)\;.
% \ee
The operator eigenvalue equations \eqref{opEeigenvalueDef1} and \eqref{optildeEeigenvalueDef1} can be translated into a Schroedinger-like equations
for the excitonic states.
